---
title: "TSA Report"
date: "2023-06-01"
output: 
  html_document:
    theme: spacelab
    highlight: tango
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      cache   = TRUE,
                      message = FALSE, 
                      warning = FALSE)
```

```{r}
getwd()
```

## 1. Data Preparation

Load necessary packages to memory:

```{r}
library(xts)
library(tidyverse)
library(quantmod)
library(lmtest)
library(forecast)
library(dygraphs)
library(vars)
library(knitr)
library(kableExtra)
```

Load some pre-defined functions:

```{r}
source("testdf.R")
```

```{r}
source("function_plot_ACF_PACF_resids.R")
```

Load Dataset:

```{r}
Data <- read.csv("TSA_2023_project_data_1.csv",header = TRUE, dec = ".")
```

## 2 Summary Data

```{r}
str(Data)
```

Notice that the first column is the Date column, which is under
"Character" type We need to transform it into "Date" type.

```{r}
Data$X <- as.Date(Data$X, 
                  format = "%d-%m-%y")
```

Until now the class is "Data.Frame" object

```{r}
class(Data) 
```

Create xts objects

```{r}
Data <- xts(Data[,-1], Data$X)
```

After creating xts objects, the class is "xts", "zoo"

```{r}
class(Data) 
```

```{r}
str(Data)
```

```{r}
Data %>% glimpse()
```

Checking for missing data

```{r}
summary(Data)
```

```{r}
head(Data,6)
```

There is no missing data for the table.

##3. Checking for Cointegration ###a/ Visualize the chart for all 10
time series.

Plot 10 Time Series together First Method

```{r}
plot(Data,  
     main = "Line Chart for 10 Time Series",  
     major.ticks = "years", 
     grid.ticks.on = "years",
     grid.ticks.lty = 3,
     legend.loc = "bottomleft",
     type="l")
```

Second Method

```{r}
Data %>%
  dygraph() %>%
  dyRangeSelector(height = 40)
```

Plot 10 Time Series separately

```{r}
plot(Data,
     main = "Line Chart for 10 Time Series",  
     major.ticks = "years", 
     grid.ticks.on = "years",
     legend.loc = "bottomleft",
     multi.panel = TRUE,
     yaxis.same = FALSE,
     type="l")
```

```{r}
plot(Data,
     main = list("Line Chart for 10 Time Series", 
                 cex = 1.5, 
                 line = 2,
                 col = "darkblue",
                 font = 2),
     major.ticks = "years", 
     grid.ticks.on = "years",
     legend.loc = "bottomleft",
     multi.panel = TRUE,
     yaxis.same = FALSE,
     type = "l")
```

```{r}
plot(Data$x1,type="l")
plot(Data$x2,type="l")
plot(Data$x2,type="l")
plot(Data$x3,type="l")
plot(Data$x4,type="l")
plot(Data$x5,type="l")
plot(Data$x6,type="l")
plot(Data$x7,type="l")
plot(Data$x8,type="l")
plot(Data$x9,type="l")
plot(Data$x10,type="l")
```

So our group decide to choose Time Series 1 and 2 for analysis.

###b. Create first difference

```{r}
Data$dx1 <- diff.xts(Data$x1)
Data$dx2 <- diff.xts(Data$x1)
Data$dx3 <- diff.xts(Data$x3)
Data$dx4 <- diff.xts(Data$x4)
Data$dx5 <- diff.xts(Data$x5)
Data$dx6 <- diff.xts(Data$x6)
Data$dx7 <- diff.xts(Data$x7)
Data$dx8 <- diff.xts(Data$x8) 
Data$dx9 <- diff.xts(Data$x9)
Data$dx10 <- diff.xts(Data$x10) 
```

Plot both variables on the graph:

```{r}
plot(Data[, 1:2],
     col = c("red", "blue"),
     major.ticks = "years", 
     grid.ticks.on = "years",
     grid.ticks.lty = 2,
     main = "Time Series X1 & X2",
     legend.loc = "topright")
```

###c Testing cointegration We perform the tests of integration order.

Testing the order of the time series 1 and its difference:

```{r}
testdf(variable = Data$x1,
       max.augmentations = 3)
```

Interprete result: The Beusch-Godfrey: Ho: There is no auto-correlation
in Residuals

```{r}
testdf(variable = Data$dx1, 
       max.augmentations = 3)
```

Since we can reject the null in the case of the first differences, we
can conclude that the Time Series X1 is integrated of order 1.

Testing the order of the time series 2 and its difference:

```{r}

testdf(variable = Data$x2,
       max.augmentations = 3)
```

```{r}

testdf(variable = Data$dx2, 
       max.augmentations = 3)
```

Since we can reject the null in the case of the first differences, we
can conclude that the Time Series X2 is integrated of order 1.

As a result, both variables has the same order 1: ∼I(1) so in the next
step we can check whether they are cointegrated or not.

###d. Estimating the Cointegrated Vector:

```{r}
model.coint <- lm(x1 ~ x2, data = Data)
```

Examine the model summary:

```{r}
summary(model.coint)
```

The model is significantly explained by x2. We further test the
stationary of the residuals:

```{r}
testdf(variable = residuals(model.coint), max.augmentations = 3)
```

The ADF test with no augmentations can be used. The result is that
non-stationarity of residuals is STRONGLY REJECTED, so residuals are
stationary, which means that x1 and x2 are cointegrated.

The cointegrating vector is [1, -29.853 , -0.713]

which defines the cointegrating relationship as: 1 \* x1 - 29.853 -
0.713 \* x2.

##4. Applying Box-Jenkins procedure for Time Series X1:

### Step 1: Model Parameters:

```{r}
par(mfrow = c(2, 1)) 
acf(Data$dx1,
    lag.max = 36, # max lag for ACF
    ylim = c(-0.1, 0.1),   # limits for the y axis - we give c(min, max)
    lwd = 5,               
    col = "dark green",
    na.action = na.pass)   # do not stop if there are missing values in the data
pacf(Data$dx1, 
     lag.max = 36, 
     lwd = 5, col = "dark green",
     na.action = na.pass)

par(mfrow = c(1, 1)) # restore the original single panel
```

The PACF shown is suggestive of an AR(5) model. So an initial candidate
model is an ARIMA(5,1,0). We also have some variations of this model:
ARIMA(5,1,1), ARIMA(4,1,0),ARIMA(3,1,0).

###Step 2 - model estimation

```{r}
arima510 <- Arima(Data$x1,  # variable	
                  order = c(5, 1, 0)  # (p,d,q) parameters	
)
```

```{r}
coeftest(arima510)
summary(arima510)
```

The above model is not included constant. Let's include constant into
the model:

```{r}
arima510_2 <- Arima(Data$x1,  # variable	
                  order = c(5, 1, 0),  # (p,d,q) parameters
                  include.constant = TRUE)  # including a constant
```

```{r}
coeftest(arima510_2)
```

Adding the constant did not change the result much, so we keep the model
without constant

###Step 3 - model diagnostics Method 1:

```{r}
plot(resid(arima510))
```

Method 2:

```{r}
tibble(
  date = index(Data),
  resid = arima510 %>% resid() %>% as.numeric()
) %>%
  ggplot(aes(date, resid)) +
  geom_line(col = "royalblue3") +
  theme_bw()
```

Lets check the ACF and the PACF of the Residual values:

Method 1:

```{r}
par(mfrow = c(2, 1)) 
acf(resid(arima510), 
    lag.max = 36,
    ylim = c(-0.1, 0.1), 
    lwd = 5, col = "dark green",
    na.action = na.pass)
pacf(resid(arima510), 
     lag.max = 36, 
     lwd = 5, col = "dark green",
     na.action = na.pass)
par(mfrow = c(1, 1))

```

Method 2:

```{r}
plot_ACF_PACF_resids(arima510)
```

# The Ljung-Box test (for a maximum of 10 lags):

```{r}
Box.test(resid(arima510), type = "Ljung-Box", lag = 10)
Box.test(resid(arima510), type = "Ljung-Box", lag = 15)
Box.test(resid(arima510), type = "Ljung-Box", lag = 20)
Box.test(resid(arima510), type = "Ljung-Box", lag = 25)
```

-\> Very large p-values: The residual is white-noise

Plot graph for Ljung-Box test:

bj_pvalues = c()

```{r}
for(i in c(1:100)){
  bj = Box.test(resid(arima510), type = "Ljung-Box", lag = i)
  bj_pvalues = append(bj_pvalues,bj$p.value)
}

plot(bj_pvalues, type='l')

abline(h = 0.05, col='red')

```

Model ARIMA(5,1,1)

```{r}

arima511 <- Arima(Data$x1,  # variable	
                    order = c(5, 1, 1),  # (p,d,q) parameters
                    include.constant = TRUE)  # including a constant
```

```{r}
coeftest(arima511)
```

ARIMA(4,1,0)

```{r}
arima410 <- Arima(Data$x1,  # variable	
                  order = c(4, 1, 0),  # (p,d,q) parameters
                  include.constant = TRUE)  # including a constant
```

```{r}
coeftest(arima410)
```

ARIMA(3,1,0).

```{r}
arima310 <- Arima(Data$x1,  # variable	
                  order = c(3, 1, 0),  # (p,d,q) parameters
                  include.constant = TRUE)  # including a constant
```

```{r}
coeftest(arima310)
```

###Step 4. Evaluate Model:

```{r}
AIC(arima510,arima510_2, arima511, arima410, arima310)
BIC(arima510, arima510_2, arima511, arima410, arima310)
```

-\> Suggestion ARIMA(5,1,0)

#Cross check with Auto Correlation:

```{r}

arima.best.AIC <- 
  auto.arima(Data$x1,
             d = 1,             # parameter d of ARIMA model
             max.p = 6,         # Maximum value of p
             max.q = 6,         # Maximum value of q
             max.order = 12,    # maximum p+q
             start.p = 1,       # Starting value of p in stepwise procedure
             start.q = 1,       # Starting value of q in stepwise procedure
             ic = "aic",        # Information criterion to be used in model selection.
             stepwise = FALSE,  # if FALSE considers all models
             allowdrift = TRUE, # include a constant
             trace = TRUE)      # show summary of all models considered
```

Return: Best model: ARIMA(4,1,2)

```{r}
arima412 <- Arima(Data$x1,  # variable	
                  order = c(4, 1, 2),  # (p,d,q) parameters
                  include.constant = TRUE)  # including a constant
```

```{r}
coeftest(arima412)
```

```{r}
AIC(arima510,arima412)
BIC(arima510,arima412)
```

Check for BIC best:

```{r}
arima.best.BIC <- 
  auto.arima(Data$x1,
             d = 1,             # parameter d of ARIMA model
             max.p = 6,         # Maximum value of p
             max.q = 6,         # Maximum value of q
             max.order = 12,    # maximum p+q
             start.p = 1,       # Starting value of p in stepwise procedure
             start.q = 1,       # Starting value of q in stepwise procedure
             ic = "bic",        # Information criterion to be used in model selection.
             stepwise = FALSE,  # if FALSE considers all models
             allowdrift = TRUE, # include a constant
             trace = TRUE)      # show summary of all models considered
```

Return: Best model: ARIMA(3,1,2)

```{r}

arima312 <- Arima(Data$x1,  # variable	
                  order = c(3, 1, 2),  # (p,d,q) parameters
                  include.constant = TRUE)  # including a constant
```

```{r}
coeftest(arima312)
```

```{r}
AIC(arima510,arima412,arima312)
BIC(arima510,arima412,arima312)
```

NNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNNN

#5. Applying Box-Jenkins procedure for Time Series X2:

## Step 1: Model Parameters:

```{r}
par(mfrow = c(2, 1)) 
acf(Data$dx2,
    lag.max = 36, # max lag for ACF
    ylim = c(-0.1, 0.1),   # limits for the y axis - we give c(min, max)
    lwd = 5,               
    col = "dark green",
    na.action = na.pass)   # do not stop if there are missing values in the data
pacf(Data$dx2, 
     lag.max = 36, 
     lwd = 5, col = "dark green",
     na.action = na.pass)

par(mfrow = c(1, 1)) # restore the original single panel

```

The PACF shown is suggestive of an AR(5) model. So an initial candidate
model is an ARIMA(5,1,0). We also have some variations of this model:
ARIMA(5,1,1), ARIMA(4,1,0),ARIMA(3,1,0).

##Step 2 - model estimation

```{r}
arima510_x2 <- Arima(Data$x2,  # variable	
                  order = c(5, 1, 0)  # (p,d,q) parameters
)	
coeftest(arima510_x2)
summary(arima510_x2)
```

The above model is not included constant. Let's include constant into
the model:

```{r}
arima510_2_x2 <- Arima(Data$x2,  # variable	
                    order = c(5, 1, 0),  # (p,d,q) parameters
                    include.constant = TRUE)  # including a constant

coeftest(arima510_2_x2)
```

Adding the constant did not change the result much, so we keep the model
without constant

##Step 3 - model diagnostics Method 1:

```{r}
plot(resid(arima510_x2))
```

The Ljung-Box test (for a maximum of 10 lags):

```{r}
Box.test(resid(arima510_x2), type = "Ljung-Box", lag = 10)
Box.test(resid(arima510_x2), type = "Ljung-Box", lag = 15)
Box.test(resid(arima510_x2), type = "Ljung-Box", lag = 20)
Box.test(resid(arima510_x2), type = "Ljung-Box", lag = 25)
```

-\> Very large p-values: The residual is white-noise

Plot graph for Ljung-Box test:

```{r}
bj_pvalues = c()

for(i in c(1:100)){
  bj = Box.test(resid(arima510_x2), type = "Ljung-Box", lag = i)
  bj_pvalues = append(bj_pvalues,bj$p.value)
}

plot(bj_pvalues, type='l')

abline(h=0.05, col='red')
```

Model ARIMA(5,1,1)

```{r}

arima511_x2 <- Arima(Data$x2,  # variable	
                  order = c(5, 1, 1),  # (p,d,q) parameters
                  include.constant = TRUE)  # including a constant
```

```{r}
coeftest(arima511_x2)
```

ARIMA(4,1,0)

```{r}
arima410_x2 <- Arima(Data$x2,  # variable	
                  order = c(4, 1, 0),  # (p,d,q) parameters
                  include.constant = TRUE)  # including a constant
```

```{r}
coeftest(arima410_x2)
```

ARIMA(3,1,0).

```{r}
arima310_x2 <- Arima(Data$x2,  # variable	
                  order = c(3, 1, 0),  # (p,d,q) parameters
                  include.constant = TRUE)  # including a constant

coeftest(arima310_x2)
```

##Step 4. Evaluate Model:

```{r}
AIC(arima510_x2,arima510_2_x2, arima511_x2, arima410_x2, arima310_x2)
BIC(arima510_x2, arima510_2_x2, arima511_x2, arima410_x2, arima310_x2)

```

-\> Suggestion model: arima510_x2 - ARIMA(5,1,0)

Cross check with Auto Correlation:

```{r}

arima.best.AIC <- 
  auto.arima(Data$x2,
             d = 1,             # parameter d of ARIMA model
             max.p = 6,         # Maximum value of p
             max.q = 6,         # Maximum value of q
             max.order = 12,    # maximum p+q
             start.p = 1,       # Starting value of p in stepwise procedure
             start.q = 1,       # Starting value of q in stepwise procedure
             ic = "aic",        # Information criterion to be used in model selection.
             stepwise = FALSE,  # if FALSE considers all models
             allowdrift = TRUE, # include a constant
             trace = TRUE)      # show summary of all models considered
```

Return: Best model: ARIMA(3,1,3)

```{r}
arima313_x2 <- Arima(Data$x2,  # variable	
                  order = c(3, 1, 3),  # (p,d,q) parameters
                  include.constant = TRUE)  # including a constant
```

```{r}
AIC(arima510_x2,arima313_x2)
BIC(arima510_x2,arima313_x2)
```

```{r}
arima.best.BIC <- 
  auto.arima(Data$x2,
             d = 1,             # parameter d of ARIMA model
             max.p = 6,         # Maximum value of p
             max.q = 6,         # Maximum value of q
             max.order = 12,    # maximum p+q
             start.p = 1,       # Starting value of p in stepwise procedure
             start.q = 1,       # Starting value of q in stepwise procedure
             ic = "bic",        # Information criterion to be used in model selection.
             stepwise = FALSE,  # if FALSE considers all models
             allowdrift = TRUE, # include a constant
             trace = TRUE)      # show summary of all models considered
```

Return: Best model: ARIMA(3,1,2)

```{r}
arima312_x2 <- Arima(Data$x2,  # variable	
                  order = c(3, 1, 2),  # (p,d,q) parameters
                  include.constant = TRUE)  # including a constant
```

```{r}
AIC(arima510_x2,arima313_x2, arima312_x2)
BIC(arima510_x2,arima313_x2,arima312_x2 )
```

##6. FORECASTING

Import Data Forecast:

```{r}
out_of_sample <-read.csv("Out_of_sample.csv",header = TRUE, dec = ".")
```

```{r}
class(out_of_sample)
```

Change Date

```{r}
out_of_sample$X <- as.Date(out_of_sample$X, 
                           format = "%d-%m-%y") 
```

Create xts objects

```{r}
out_of_sample <- xts(out_of_sample[,-1], out_of_sample$X)
```

Assign forecast X1 to object oos_x1

```{r}
oos_x1 <- out_of_sample$x1
oos_x1

```

#FORECAST X1

```{r}
tail(Data, 12)
```

```{r}
forecasts_x1 <- forecast(arima510, # model for prediction
                      h = 30) # how many periods outside the sample

forecasts_x1
```

#Extract the forecast points

```{r}
forecasts_x1$mean
class(forecasts_x1$mean)
```

It is a ts object, not xts! However, the xts objects are more convenient
and modern. In terms of plotting the real data and forecast data to
compare.

```{r}
forecasts_x1$lower
forecasts_x1$upper
```

Create xts object: (We use the second column (95% confidence interval))

```{r}
forecasts_x1_data <- data.frame(f_mean = as.numeric(forecasts_x1$mean),
                             f_lower = as.numeric(forecasts_x1$lower[, 2]),
                             f_upper = as.numeric(forecasts_x1$upper[, 2]))
```

```{r}
head(forecasts_x1_data,30)
```

Adding real value X1 below the current Dataset:

```{r}
Data_x1 <- rbind(Data[, "x1"], oos_x1)
tail(Data_x1, n = 40)
```

Turn Forecast with Index into Forecast with Date

```{r}
forecasts_x1_xts <- xts(forecasts_x1_data,
                     order.by = index(oos_x1))
forecasts_x1_xts
```

Merge it together with the original data

```{r}
Data_x1_combined <- merge(Data_x1, forecasts_x1_xts)
head(Data_x1_combined)
tail(Data_x1_combined,40)
```

Plot Chart

```{r}
plot(Data_x1_combined [, c("x1", "f_mean", "f_lower", "f_upper")], 
     major.ticks = "years", 
     grid.ticks.on = "years",
     grid.ticks.lty = 3,
     main = "30 day forecast of x1",
     col = c("black", "blue", "red", "red"))
```

```{r}
plot(Data_x1_combined ["2020-11/", c("x1", "f_mean", "f_lower", "f_upper")], 
     major.ticks = "years", 
     grid.ticks.on = "years",
     grid.ticks.lty = 3,
     main = "30 day forecast of x1",
     col = c("black", "blue", "red", "red"))
```

Extract Data for Evaluating the Forecast:

```{r}
Data_x1_Eva <- tail(Data_x1_combined, 30)  
Data_x1_Eva

```

```{r}
Data_x1_Eva$mae   <-  abs(Data_x1_Eva$x1 - Data_x1_Eva$f_mean)
Data_x1_Eva$mse   <-  (Data_x1_Eva$x1 - Data_x1_Eva$f_mean) ^ 2
Data_x1_Eva$mape  <-  abs((Data_x1_Eva$x1 - Data_x1_Eva$f_mean)/Data_x1_Eva$x1)
Data_x1_Eva$amape <-  abs((Data_x1_Eva$x1 - Data_x1_Eva$f_mean)/(Data_x1_Eva$x1 + Data_x1_Eva$f_mean))
Data_x1_Eva
```

```{r}
colMeans(Data_x1_Eva[, c("mae", "mse", "mape", "amape")])
apply(Data_x1_Eva[, c("mae", "mse", "mape", "amape")], 2, FUN = median)
```

########################################################################################## 

#FORECAST X2

Assign forecast X2 to object oos_x2

```{r}
oos_x2 <- out_of_sample$x2

oos_x2
```

```{r}
tail(Data, 12)

forecasts_x2 <- forecast(arima510_x2, # model for prediction
                         h = 30) # how many periods outside the sample

forecasts_x2
```

Extract the forecast points

```{r}
forecasts_x2$mean
class(forecasts_x2$mean)
```

It is a ts object, not xts! However, the xts objects are more convenient
and modern. In terms of plotting the real data and forecast data to
compare.

```{r}
forecasts_x2$lower
forecasts_x2$upper
```

Create xts object: (We use the second column (95% confidence interval))

```{r}
forecasts_x2_data <- data.frame(f_mean = as.numeric(forecasts_x2$mean),
                                f_lower = as.numeric(forecasts_x2$lower[, 2]),
                                f_upper = as.numeric(forecasts_x2$upper[, 2]))
```

```{r}
head(forecasts_x2_data,30)
```

Adding real value X1 below the current Dataset:

```{r}
Data_x2 <- rbind(Data[, "x2"], oos_x2)
tail(Data_x2, n = 40)
```

Turn Forecast with Index into Forecast with Date

```{r}
forecasts_x2_xts <- xts(forecasts_x2_data,
                        order.by = index(oos_x2))
forecasts_x2_xts
```

Merge it together with the original data

```{r}
Data_x2_combined <- merge(Data_x2, forecasts_x2_xts)
head(Data_x2_combined)
tail(Data_x2_combined,40)
```

Plot Chart

```{r}
plot(Data_x2_combined ["2020-11/", c("x2", "f_mean", "f_lower", "f_upper")], 
     major.ticks = "years", 
     grid.ticks.on = "years",
     grid.ticks.lty = 3,
     main = "30 day forecast of x2",
     col = c("black", "blue", "red", "red"))
```

Extract Data for Evaluating the Forecast:

```{r}
Data_x2_Eva <- tail(Data_x2_combined, 30)  
Data_x2_Eva
```

```{r}
Data_x2_Eva$mae   <-  abs(Data_x2_Eva$x2 - Data_x2_Eva$f_mean)
Data_x2_Eva$mse   <-  (Data_x2_Eva$x2 - Data_x2_Eva$f_mean) ^ 2
Data_x2_Eva$mape  <-  abs((Data_x2_Eva$x2 - Data_x2_Eva$f_mean)/Data_x2_Eva$x2)
Data_x2_Eva$amape <-  abs((Data_x2_Eva$x2 - Data_x2_Eva$f_mean)/(Data_x2_Eva$x2 + Data_x2_Eva$f_mean))
Data_x2_Eva
```

```{r}
colMeans(Data_x2_Eva[, c("mae", "mse", "mape", "amape")])
apply(Data_x2_Eva[, c("mae", "mse", "mape", "amape")], 2, FUN = median)
```
